Preprocessing

### 2024-08-13 nbs/guides/preprocessing/preprocessing.qmd
fix(nbs): plot mitochondrial and ribosomal percentage histograms in preprocessing guide
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

if PYROVELOCITY_DATA_SUBSET:
    adata = adata[
        np.random.choice(
            adata.obs.index,
            size=300,
            replace=False,
        ),
        :,
    ].copy()
subset_anndata_representation = anndata_string(adata)
print_string_diff(
    text1=initial_data_state_representation,
    text2=subset_anndata_representation,
    diff_title="Downsampled data diff",
)
```

@@

@@

adata.var["mt"] = adata.var_names.str.startswith("MT-" or "Mt-" or "mt-")
adata.var["ribo"] = adata.var_names.str.startswith(
    "RPS" or "Rps" or "rps" or "RPL" or "Rpl" or "rpl"
)

@@

@@

chart = plot_spliced_unspliced_histogram(
    adata=adata,
    spliced_layer="raw_spliced",
    unspliced_layer="raw_unspliced",
    min_count=3,
    max_count=200,
    title="",
    plot_width=700,
    plot_height=700,
)
chart.save("us_histogram.pdf")
chart
```

>>>>

if PYROVELOCITY_DATA_SUBSET:
    adata = adata[
        np.random.choice(
            adata.obs.index,
            size=300,
            replace=False,
        ),
        :,
    ].copy()

subset_anndata_representation = anndata_string(adata)
print_string_diff(
    text1=initial_data_state_representation,
    text2=subset_anndata_representation,
    diff_title="Downsampled data diff",
)
```

@@

Prior to filtering any genes or cells we compute the quality control metrics 
including the percentage of mitochondrial and ribosomal counts in each cell.

@@

adata.var["mt"] = adata.var_names.str.startswith(("MT-", "Mt-", "mt-"))
adata.var["ribo"] = adata.var_names.str.startswith(
    ("RPS", "Rps", "rps", "RPL", "Rpl", "rpl")
)

@@

We plot the percentage of mitochondrial and ribosomal counts in each cell

```{python}
# | label: pre-filtration-plot-mitochondrial
import seaborn as sns

ax = sns.histplot(
    adata.obs,
    x="pct_counts_mt",
    color="#ff6a14",
)
ax.set_xlabel("mitochondrial counts (%)")
```

```{python}
# | label: pre-filtration-plot-ribosomal
ax = sns.histplot(
    adata.obs,
    x="pct_counts_ribo",
    color="#ff6a14",
)
ax.set_xlabel("ribosomal counts (%)")
```

as well as the number of genes with at least one count in each cell and the
total counts in each cell.

@@

chart = plot_spliced_unspliced_histogram(
    adata=adata,
    spliced_layer="raw_spliced",
    unspliced_layer="raw_unspliced",
    min_count=3,
    max_count=200,
)
chart.save("us_histogram.pdf")
chart
```
